# Data transformation

```{r, echo=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(countrycode)
```

We aim at merging all data into one tidy master data frame within which each observation will correspond to one country and one year.

## Selecting the time range

We hope to keep a sufficiently wide range so as to observe clear change in energy consumption behavior across time. 


```{r}
# Load all data sets
df_population_wide <- read.csv("data/population.csv", skip = 4, check.names = FALSE)
df_GDP_wide <- read.csv("data/worldbank_GDP.csv", skip = 4, check.names = FALSE)
df_HDI_wide <- read.csv("data/HDI.csv", skip = 5, check.names = FALSE)
df_life_wide <- read.csv("data/life_expectancy.csv", skip = 4, check.names = FALSE)
df_schooling_wide <- read.csv("data/un_years_of_schooling.csv", skip = 6, check.names = FALSE)
df_energy_long <- read.csv("data/owid_energy_data.csv", header = TRUE, stringsAsFactors = FALSE)
df_country_info <- read.csv("data/worldbank_country_info.csv", header = TRUE, stringsAsFactors = FALSE, encoding = "UTF-8")
```

```{r}
year_availability <- matrix(NaN, 6, 2)
colnames(year_availability) <- c("Min year", "Max year")
rownames(year_availability) <- c("Energy consumption" , "Years of schooling" , "Life expectancy" , "HDI" , "GDP", "Population")

year_availability[6,] <- c(1960, 2020)
year_availability[5,] <- c(1960, 2020)
year_availability[4,] <- c(1990, 2019)
year_availability[3,] <- c(1960, 2020)
year_availability[2,] <- c(1990, 2019)
year_availability[1,] <- c(1900, 2020)

## ncol and nrow
nr <- nrow(year_availability)
nc <- ncol(year_availability)

## separation between bars
sepr <- 0.2

op <- par(mar = c(5,9,4,2) + 0.1)

plot(c(1900,2020), c(1,nr), type="n",
     ylim=c(1-(1-sepr)/2-sepr,
            nr+(1-sepr)/2+sepr),
     ylab="", yaxt="n",
     xlab="", main = "Time range for each data set")
grid(ny = NA)

## plot bars
for(i in 1:nr) {
    lo <- year_availability[i,1]
    hi <- year_availability[i,2]
    polygon(x=c(lo, lo, hi, hi),
            y=c(i-(1-sepr)/2, i+(1-sepr)/2,
                i+(1-sepr)/2, i-(1-sepr)/2), col = "#6495ED")
}

## create y-axis
axis(2, at=1:nr, las=2, xpd=NA, labels=rownames(year_availability)) 
legend("topleft", fill="#6495ED", bty="n", legend="year is recorded")
## add center line
abline(v=0, lty=2)
```

We choose to restrict our multivariate analysis to the narrowest time range. This is exactly the intersection of all ranges : **1990-2019**.

We note that this plot **does not tell us anything about missing values**. Indeed, a given observation might be missing although within the time range of the data set.

* * *

## Selecting the variables

We hope to be able to analyze and understand the cross-correlation between energy consumption and human development. To do so, we figured we would need the following:

- data indicators for **HDI and all its major components**, namely: `GDP`, `life_expectancy`, `years_of_schooling`;
- figures for **consumption per energy type**;
- population.


```{r}
# except from year and country
variable <- c("ISO_code", "HDI", "population", 
             "years_of_schooling", "life_expectancy", "GDP", "fossil_fuel", "hydro", "low_carb", 
               "nuclear", "other_renew", "renew", "solar", "wind", 
               "coal", "gas", "oil", "biofuel", "primary")
dataset <- c("worldbank_country_info", "HDI", "population", "un_years_of_schooling", "life_expectancy",
             "worldbank_gdp", "owid_energy_data", "owid_energy_data", "owid_energy_data", "owid_energy_data", "owid_energy_data", "owid_energy_data", "owid_energy_data", "owid_energy_data", "owid_energy_data", "owid_energy_data", "owid_energy_data", "owid_energy_data", "owid_energy_data")

chosen_variables <- data.frame(variable, dataset)

knitr::kable(chosen_variables, "html")
```

```{r, warning=FALSE}
# Choose years
relevant_years <- as.character(1990:2019)

df_energy_long <- df_energy_long %>% 
  select(-contains(c("_per_capita", "_pct", "_change", "_production", "_share"))) %>%
  subset(year %in% relevant_years) %>%
  rename("ISO_code" = "iso_code") %>%
  subset(ISO_code != "") %>%
  rename("fossil_fuel" = "fossil_fuel_consumption", 
         "hydro" = "hydro_consumption", 
         "low_carb" = "low_carbon_consumption", 
         "nuclear" = "nuclear_consumption", 
         "other_renew" = "other_renewable_consumption", 
         "renew" = "renewables_consumption", 
         "solar" = "solar_consumption", 
         "wind" = "wind_consumption", 
         "coal" = "coal_consumption", 
         "gas" = "gas_consumption", 
         "oil" = "oil_consumption", 
         "biofuel" = "biofuel_consumption", 
         "primary" = "primary_energy_consumption") %>%
  select("ISO_code", "year", "fossil_fuel", "hydro", "low_carb", "nuclear", "other_renew", "renew", 
         "solar", "wind", "coal", "gas", "oil", "biofuel", "primary")

df_energy_ISO_with_no_missing_year <- df_energy_long %>%
  select("ISO_code", "year") %>%
  group_by(ISO_code) %>%
  summarize(count = n()) %>%
  filter(count == 2019 - 1990 + 1) %>%
  pull(ISO_code)

df_energy_long <- df_energy_long %>%
  filter(ISO_code %in% df_energy_ISO_with_no_missing_year)



df_GDP_wide <- df_GDP_wide %>%
  select("Country Name", "Country Code", relevant_years) %>%
  rename("country" = "Country Name", "ISO_code" = "Country Code") %>%
  mutate_at(relevant_years, as.numeric) %>%
  select("ISO_code", relevant_years)
  


df_life_wide <- df_life_wide %>%
  select("Country Name", "Country Code", relevant_years) %>%
  rename("country" = "Country Name", "ISO_code" = "Country Code") %>%
  mutate_at(relevant_years, as.numeric) %>%
  select("ISO_code", relevant_years)


df_schooling_wide <- df_schooling_wide %>%
  select("Country", relevant_years) %>%
  mutate("ISO_code" = countrycode(Country, origin = "country.name", destination = "iso3c")) %>%
  mutate_at(relevant_years, as.numeric) %>%
  select("ISO_code", relevant_years)


df_population_wide <- df_population_wide %>%
  select("Country Name", "Country Code", relevant_years) %>%
  rename("country" = "Country Name", "ISO_code" = "Country Code") %>%
  mutate_at(relevant_years, as.numeric) %>%
  select("ISO_code", relevant_years)


df_HDI_wide <- df_HDI_wide %>%
  select("Country", relevant_years) %>%
  mutate("ISO_code" = countrycode(Country, origin = "country.name", destination = "iso3c")) %>%
  mutate_at(relevant_years, as.numeric) %>%
  select("ISO_code", relevant_years)


df_country_info <- as.data.frame(df_country_info) %>% 
  rename("ISO_code" = "X.U.FEFF.iso_code", "region" = "Region", "income_group" = "IncomeGroup", "special_notes" = "SpecialNotes")


# retrieve common countries across data sets based on ISO code
common_ISO_codes <- Reduce(intersect, list(df_energy_long$ISO_code,df_life_wide$ISO_code, df_schooling_wide$ISO_code, 
                                           df_GDP_wide$ISO_code, df_population_wide$ISO_code, df_HDI_wide$ISO_code))

# filter data sets on common ISO codes
df_energy_long <- df_energy_long %>% filter(ISO_code %in% common_ISO_codes)
df_GDP_wide <- df_GDP_wide %>% filter(ISO_code %in% common_ISO_codes)
df_life_wide <- df_life_wide %>% filter(ISO_code %in% common_ISO_codes)
df_schooling_wide <- df_schooling_wide %>% filter(ISO_code %in% common_ISO_codes)
df_population_wide <- df_population_wide %>% filter(ISO_code %in% common_ISO_codes)
df_HDI_wide <- df_HDI_wide %>% filter(ISO_code %in% common_ISO_codes)
```


```{r}
# Convert data frames to long format
df_GDP_long <- df_GDP_wide %>% 
  pivot_longer(cols = -c("ISO_code"), names_to = "year", values_to = "GDP")
df_schooling_long <- df_schooling_wide %>% 
  pivot_longer(cols = -c("ISO_code"), names_to = "year", values_to = "years_of_schooling") %>%
  mutate_at("year", as.numeric)
df_life_long <- df_life_wide %>% pivot_longer(cols = -c("ISO_code"), names_to = "year", values_to = "life_expectancy")
df_population_long <- df_population_wide %>% 
  pivot_longer(cols = -c("ISO_code"), names_to = "year", values_to = "population") 
df_HDI_long <- df_HDI_wide %>% pivot_longer(cols = -c("ISO_code"), names_to = "year", values_to = "HDI")
```


```{r}
# display number of rows for each dataframe
nrow(df_energy_long)
nrow(df_GDP_long)
nrow(df_life_long)
nrow(df_schooling_long)
nrow(df_population_long)
nrow(df_HDI_long)
```
```{r}
# merge data frames into one master data frame
master_df <- merge(df_energy_long, merge(df_GDP_long, merge(df_life_long, merge(df_schooling_long, merge(df_population_long, df_HDI_long,
                by = c("ISO_code", "year")), by = c("ISO_code", "year")), by = c("ISO_code", "year")), 
                by = c("ISO_code", "year")), by = c("ISO_code", "year"))

# Add one variable for country name and reorder columns
master_df <- master_df %>%
  mutate(country = countrycode(ISO_code, origin = "iso3c", destination = "country.name")) %>%
  select(ISO_code, country, year, population, GDP, HDI, life_expectancy, years_of_schooling, everything())
```

```{r}
# write master_df to csv
write.csv(master_df,"data/master.csv", row.names = FALSE)
```

#TODOS:
- What countries are missing from the data set that exist in real life
- How many regions maximum compared to final dataset
- Make data cleaning steps more explicit
  - removing variables
  - limiting time frame
  - merge and exclude duplicates




