# Results

```{r, echo=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
library(patchwork)
library(Lock5withR)
library(ggpubr)
library(stringr)
source("R/plot_missing_patterns.R")
```


```{r echo=FALSE, warning=FALSE}
# Load Data
master_df <- read.csv(file = "data/master.csv")
```

## Jannik

## Is an energy mix oriented towards renewable sources associated with good human development indicators? 

```{r}
df_graph <- master_df[master_df$year==2019,]
ggplot(df_graph, aes(x=primary/population*1000000000, y=years_of_schooling)) +
  geom_point(aes(color = year)) +
  geom_smooth(method="lm") +
  theme_light() + 
  geom_text(
    label=df_graph$country, 
    nudge_x = 0, nudge_y = 0.4,
    size = 3,
    check_overlap = T
  ) +
  scale_x_log10()
```

```{r}
selected_cols <- c("renewables_share_elec", "low_carbon_share_energy", "hdi_rank", "schooling_years")
selected_ids <- match(selected_cols, names(df_energy_and_schooling))

pairs(df_energy_and_schooling[df_energy_and_schooling, selected_ids])
```

## What characterises the countries that have the most renewable energy mix? 

### Ideas:
- relationship between HDI and different energy metrics -> per capita, coal, renewable, etc.
- outlier analysis
- filter on GDP or population to make trends appear
- cluster countries by energy mix



On the other hand, how does the production of coal and gas energy leading to more pollution might foster and hinder human development at the same time?
- countries might consume other energy than they produce
- 


## Antonin

```{r, fig.height=8}
master_df <- read.csv(file = "data/master.csv")
plot_missing_parameters(master_df, use_counts = FALSE, max_100 = TRUE, 
                        ticks_angle = 45, ticks_size = 9, complete_cases_size = 3.5, complete_cases_hjust = 0.1, y_ticks_size = 9)
```



