# Interactive component

```{r, echo=FALSE, warning=FALSE}
library(dplyr)
library(tidyverse)
```

```{r echo=FALSE, warning=FALSE}
# Load Data
master_df <- read.csv(file = "data/master.csv")
distinct_energy_vars <- c("hydro", "nuclear", "solar", "wind", "coal", "gas", "oil", "biofuel", "other_renew")

left_df <- master_df %>%
  select(c("ISO_code", "year", "HDI", "solar", "biofuel", "wind", "coal", "nuclear", "other_renew", "gas", "oil", "hydro")) %>%
  na.omit()

right_df <- left_df[left_df$year==2019,] %>% select(c("ISO_code", "HDI"))

interactive_df <- left_join(left_df, right_df, by="ISO_code") %>%
  mutate(HDI = HDI.y) %>%
  select(-c("HDI.x", "HDI.y")) %>%
  mutate(HDI = case_when(HDI <= 0.549 ~ 'Low',
                         (HDI <= 0.699 & HDI > 0.549) ~ 'Medium',
                         (HDI <= 0.799 & HDI > 0.699) ~ 'High',
                         (HDI <= 1 & HDI > 0.799) ~ 'Very high')) %>%
  select(-"ISO_code") %>%
  group_by(year, HDI) %>%
  summarise_all(list(sum)) %>%
  rename(Year="year", Solar="solar", Biofuel="biofuel", Wind="wind",
         Coal="coal", Nuclear="nuclear", 
         "Other_renewables"="other_renew", Gas="gas", Oil="oil",
          Hydropower="hydro")

write.csv(interactive_df,"data/interactive6.csv", row.names = FALSE)
```


## Interactive visualization on the energy mix by HDI level {.unlisted .unnumbered}

You can do the following:

- Selecting only a part of the y-axis to identify the smaller categories
- Hover over each element of the legend to highlight the variable in the graph
- Select countries of which HDI level you want to examine

<style>
  body {
    font-family: Verdana, Arial, sans-serif;
  }

  h1 {
    font-size: 1.2em;
    margin: 60px 0 5px 0;
  }

  svg {
    width: 1200px;
    margin-left: auto;
    margin-right: auto;
  }

  b {
    margin-left: 50px;
  }

  .wrapper {
    width: 800px;
    margin-left: 100px;
  }

  .HDI_category {
    width: 50px;
    text-align: center;
  }
</style>


<!-- Load d3.js -->
<script src="https://d3js.org/d3.v7.js"></script>

<!-- Create a div where the graph will take place -->



  <div id="my_dataviz"></div>
  <!-- Initialize a select button -->
  <b>Select the wanted HDI category: </b>
  <select id="HDI_category"></select>
</body>
<script>
  // set the dimensions and margins of the graph
  const margin = { top: 60, right: 230, bottom: 20, left: 50 },
    width = 800 - margin.left - margin.right,
    height = 500 - margin.top - margin.bottom;

  // append the svg object to the body of the page
  const svg = d3
    .select("#my_dataviz")
    .append("svg")
    .attr("width", width + margin.left + margin.right)
    .attr("height", height + margin.top + margin.bottom)
    .append("g")
    .attr("transform", `translate(${margin.left}, ${margin.top})`);

  // Parse the Data
  d3.csv(
    "https://raw.githubusercontent.com/AntoninVidon/energyhumandevelopment/main/data/interactive7.csv"
  ).then(function (data) {
    // List of groups (here I have one group per column)
    const allGroup = new Set(data.map((d) => d.HDI));

    // add the options to the button
    d3.select("#HDI_category")
      .selectAll("myOptions")
      .data(allGroup)
      .enter()
      .append("option")
      .text(function (d) {
        return d;
      }) // text showed in the menu
      .attr("value", function (d) {
        return d;
      }); // corresponding value returned by the button

    // store old data
    old_data = data
    data = old_data.filter(function (d) {
      return d.HDI == "High";
    });

    //////////
    // GENERAL //
    //////////

    // List of groups = header of the csv files
    //const keys = data.columns.slice(1)
    const keys = [
      "Coal",
      "Oil",
      "Gas",
      "Hydropower",
      "Nuclear",
      "Solar",
      "Biofuel",
      "Wind",
      "Other_renewables",
    ];

    // color palette
    const color = d3
      .scaleOrdinal()
      .domain(keys)
      .range([
        "#645f55",
        "black",
        "#b18781",
        "#1e7991",
        "#c0b879",
        "#eabc35",
        "#8d9e2b",
        "#7cabcb",
        "#d0d890",
      ]);

    //stack the data?
    const stackedData = d3.stack().keys(keys)(data);

    //////////
    // AXIS //
    //////////

    // Add X axis
    const x = d3
      .scaleLinear()
      .domain(
        d3.extent(data, function (d) {
          return d.Year;
        })
      )
      .range([0, width]);
    const xAxis = svg
      .append("g")
      .attr("transform", `translate(0, ${height})`)
      .call(d3.axisBottom(x).ticks(5).tickFormat(d3.format("d")));

    // Add X axis label:
    svg
      .append("text")
      .attr("text-anchor", "end")
      .attr("x", width)
      .attr("y", height + 40)
      .text("Time (year)");

    // Add Y axis label:
    svg
      .append("text")
      .attr("text-anchor", "end")
      .attr("x", 0)
      .attr("y", -20)
      .text("Energy consumption split for the selected HDI range (TWh)")
      .attr("text-anchor", "start");

    // Add Y axis
    const y = d3
      .scaleLinear()
      .domain([
        0,
        d3.max(data, function (d) {
          return (
            +d.Hydropower +
            +d.Nuclear +
            +d.Solar +
            +d.Wind +
            +d.Coal +
            +d.Gas +
            +d.Oil +
            +d.Biofuel +
            +d.Other_renewables
          );
        }) * 1.1,
      ])
      .range([height, 0]);
    const yAxis = svg
      .append("g")
      .attr("transform", `translate(0, 0)`)
      .call(d3.axisLeft(y).ticks(5));

    //////////
    // BRUSHING AND CHART //
    //////////

    // Add a clipPath: everything out of this area won't be drawn.
    const clip = svg
      .append("defs")
      .append("svg:clipPath")
      .attr("id", "clip")
      .append("svg:rect")
      .attr("width", width)
      .attr("height", height)
      .attr("x", 0)
      .attr("y", 0);

    // Add brushing
    const brush = d3
      .brushY() // Add the brush feature using the d3.brush function
      .extent([
        [0, 0],
        [width, height],
      ]) // initialise the brush area: start at 0,0 and finishes at width,height: it means I select the whole graph area
      .on("end", updateChart); // Each time the brush selection changes, trigger the 'updateChart' function

    // Create the scatter variable: where both the circles and the brush take place
    const areaChart = svg.append("g").attr("clip-path", "url(#clip)");

    // Area generator
    const area = d3.area()
      .x(function (d) {
        return x(d.data.Year);
      })
      .y0(function (d) {
        return y(d[0]);
      })
      .y1(function (d) {
        return y(d[1]);
      });

    // Show the areas
    areaChart
      .selectAll("mylayers")
      .data(stackedData)
      .join("path")
      .attr("class", function (d) {
        return "myArea " + d.key;
      })
      .style("fill", function (d) {
        return color(d.key);
      })
      .attr("d", area);

    // Add the brushing
    areaChart.append("g").attr("class", "brush").call(brush);

    let idleTimeout;
    function idled() {
      idleTimeout = null;
    }

    // A function that update the chart for given boundaries
    function updateChart(event, d) {
      extent = event.selection;

      // If no selection, back to initial coordinate. Otherwise, update X axis domain
      if (!extent) {
        if (!idleTimeout) return (idleTimeout = setTimeout(idled, 350)); // This allows to wait a little bit
        y.domain(
          d3.extent(data, function (d) {
            return d.Year;
          })
        );
      } else {
        y.domain([y.invert(extent[1]), y.invert(extent[0])]);
        areaChart.select(".brush").call(brush.move, null); // This remove the grey brush area as soon as the selection has been done
      }

      // Update axis and area position
      yAxis.transition().duration(1000).call(d3.axisLeft(y).ticks(5));
      areaChart.selectAll("path").transition().duration(1000).attr("d", area);
    }

    // A function that update the chart
    function update(selectedGroup) {
      //stack the data?

      // get new data
      data = old_data.filter(function (d) {
        return d.HDI == selectedGroup;
      });

      const stackedData = d3.stack().keys(keys)(data);

      // set new y domain according to the new data
      y.domain([
        0,
        d3.max(data, function (d) {
          return (
            +d.Hydropower +
            +d.Nuclear +
            +d.Solar +
            +d.Wind +
            +d.Coal +
            +d.Gas +
            +d.Oil +
            +d.Biofuel +
            +d.Other_renewables
          );
        }),
      ]);

      // update y axis
      yAxis.transition().duration(1000).call(d3.axisLeft(y).ticks(5));
      // update areaChart

      console.log(data);
      areaChart.selectAll("path").remove();

      // Show the areas
      areaChart
        .selectAll("mylayers")
        .data(stackedData)
        .join("path")
        .transition()
        .duration(1000)
        .attr("class", function (d) {
          return "myArea " + d.key;
        })
        .style("fill", function (d) {
          return color(d.key);
        })
        .attr("d", area);
    }

    //////////
    // HIGHLIGHT GROUP //
    //////////

    // What to do when one group is hovered
    const highlight = function (event, d) {
      // reduce opacity of all groups
      d3.selectAll(".myArea").style("opacity", 0.1);
      // expect the one that is hovered
      d3.select("." + d).style("opacity", 1);
    };

    // And when it is not hovered anymore
    const noHighlight = function (event, d) {
      d3.selectAll(".myArea").style("opacity", 1);
    };

    //////////
    // LEGEND //
    //////////

    // Add one dot in the legend for each name.
    const size = 20;

    svg
      .selectAll("myrect")
      .data(keys)
      .join("rect")
      .attr("x", width + 10)
      .attr("y", function (d, i) {
        return 10 + i * (size + 5);
      }) // 100 is where the first dot appears. 25 is the distance between dots
      .attr("width", size)
      .attr("height", size)
      .style("fill", function (d) {
        return color(d);
      })
      .on("mouseover", highlight)
      .on("mouseleave", noHighlight);

    // Add one dot in the legend for each name.


    svg
      .selectAll("mylabels")
      .data(keys)
      .join("text")
      .attr("x", width + 10 + size * 1.2)
      .attr("y", function (d, i) {
        return 10 + i * (size + 5) + size / 2;
      }) // 100 is where the first dot appears. 25 is the distance between dots
      .style("fill", function (d) {
        return color(d);
      })
      .text(function (d) {
        return d;
      })
      .attr("text-anchor", "left")
      .style("alignment-baseline", "middle")
      .on("mouseover", highlight)
      .on("mouseleave", noHighlight);

    // When the button is changed, run the updateChart function
    d3.select("#HDI_category").on("change", function (event, d) {
      // recover the option that has been chosen
      const selectedOption = d3.select(this).property("value");
      // run the updateChart function with this selected option
      update(selectedOption);
    });
  });
</script>